---
title: Text Cleaning
author: Kevon  Davenport
date: '2021-03-29'
slug: text-cleaning
categories:
  - R
tags:
  - homework
---
Loading the tidyverse and ggalt library into my notebook.
```{r}
library(tidyverse)
library(ggalt)
```
Creating a data frame called "logs" which reads and holds my data.
```{r}
logs <- read.csv("badfootballlogs19.csv")
```

This block separates the initial "Result" column into 3 different columns: "Outcome", "TeamScore", and "OpponentScore". 
```{r}
newlogs <- logs %>% 
  separate(Result, into=c("Outcome", "Score"), sep="\\(")  %>% mutate(Score=gsub("\\).*","",Score)) %>% separate(Score, into = c("TeamScore","OpponentScore"), sep = "-")
```

This block forces R to read out  Team and Opponent scores as numbers instead of text.
```{r}
newlogs %>% mutate(TeamScore = as.numeric(TeamScore), OpponentScore = as.numeric(OpponentScore)) %>% mutate(Differential = OpponentScore - TeamScore) -> Data
```
This block creates a new dataframe for games with a differential of greater than 65.
```{r}
worstgames <- Data %>% filter(Differential > 65)
```
This block graphs my Differential data & my worst games data and graphs it. It also encircles my worst games and highlights them red.
```{r}
ggplot() + 
  geom_point(data=Data, aes(x=TeamScore, y=OpponentScore), color="grey") +
geom_point(data=worstgames, aes(x=TeamScore, y=OpponentScore), color="red") +
geom_encircle(data=worstgames, aes(x=TeamScore, y=OpponentScore), s_shape=.001, expand=0.01, colour="red") + geom_text(aes(x=40, y=85, label="Differential"))  + labs(title="Does Offense or Defense Win Games? ", subtitle="Here's the differential of game scores. ", x="Team Score", y="Opponent Score", caption = "Source: Matt Waite | By Kevon Davenport") +  theme_minimal() + 
  theme(
    plot.title = element_text(size = 16, face = "bold"),
    axis.title = element_text(size = 8), 
    plot.subtitle = element_text(size=10), 
    panel.grid.minor = element_blank()
    )
```




